---
title: "Final Project"
author: "Charley Teltschik"
subtitle: MGSC 310 Final Project Code
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library(knitr)

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = FALSE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```


<!-- ### start answering your problem set here -->
<!-- You may export your homework in either html or pdf, with the former usually being easier. 
     To export or compile your Rmd file: click above on 'Knit' then 'Knit to HTML' -->
<!-- Be sure to submit both your .Rmd file and the compiled .html or .pdf file for full credit -->


```{r setup-3}

# load all your libraries in this chunk 
library('readr')
library('ggplot2')
library('dplyr')
library('tidyverse')
library('rsample')
library('forcats')
library('randomForest')
library('glmnet')
library('glmnetUtils')
# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```

## DATA CLEANING

READING IN DATA:

```{r}

data2019 <- read_csv("/Users/charleyteltschik/Documents/MGSC_310/datasets/2019.csv")
data2020 <- read_csv("/Users/charleyteltschik/Documents/MGSC_310/datasets/2020.csv")

```


DATA CLEANING:

```{r}

data2019 <- data2019 %>% rename(`Country` = `Country or region`,
                                `Happiness score` = Score)
data2020 <- data2020 %>% rename(`Country` = `Country name`,
                                `Happiness score` = `Ladder score`)

names(data2019)
names(data2020)

```


## DATA 2019

THIS CODE DEMONSTRATES DATA WITH RECORDED HAPPINESS SCORES AS IS

```{r}

data_happy7 <- data2019 %>% filter(`Happiness score` >= 7.0) %>% select(Country | `Happiness score`)
print(data_happy7)

```

```{r}

ggplot(data_happy7, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Countries w/Happiness Score >= 7 (2019)")

```

```{r}
data_happy6 <- data2019 %>% filter(`Happiness score` >= 6.0, `Happiness score` < 7.0) %>% 
  select(Country | `Happiness score`)
print(data_happy6)

```

```{r}
ggplot(data_happy6, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 6 & < 7 (2019)")

```

```{r}
data_happy5 <- data2019 %>% filter(`Happiness score` >= 5.0, `Happiness score` < 6.0) %>% 
  select(Country | `Happiness score`)
print(data_happy5)

```

```{r}
ggplot(data_happy5, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 5 & < 6 (2019)")

```

```{r}
data_happy4 <- data2019 %>% filter(`Happiness score` >= 4.0, `Happiness score` < 5.0) %>% 
  select(Country | `Happiness score`)
print(data_happy4)

```

```{r}
ggplot(data_happy4, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 4 & < 5 (2019)")

```

```{r}
data_happy3 <- data2019 %>% filter(`Happiness score` >= 3.0, `Happiness score` < 4.0) %>% 
  select(Country | `Happiness score`)
print(data_happy3)

```

```{r}
ggplot(data_happy3, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 3 & < 4 (2019)")

```

```{r}
data_happy2 <- data2019 %>% filter(`Happiness score` >= 2.0, `Happiness score` < 3.0) %>% 
  select(Country | `Happiness score`)
print(data_happy2)

```

```{r}
ggplot(data_happy2, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 2 & < 3 (2019)")

```

DESCRIPTIVE PLOT 2019 W/FACET WRAP

```{r}
data_list <- list(`Scores 7+` = data_happy7, `7 > Scores > 6` = data_happy6, `6 > Scores > 5` = data_happy5,
                  `4 > Scores > 5` = data_happy4, `3 > Scores > 4` = data_happy3, `2 > Scores > 3` = data_happy2)
combined_data <- bind_rows(data_list, .id = "dataset")
ggplot(combined_data, aes(x = Country, y = `Happiness score`)) +
  geom_point() +
  labs(title = "All Happiness Scores 2019", x = "Country", y = "Happiness Score (out of 10)") +
  facet_wrap(~ dataset)

```


# SPLITTING DATA INTO TRAIN AND TEST SETS

```{r}

data2019_split <- initial_split(data2019, prop = 0.75)
data2019_train <- training(data2019_split)
data2019_test <- testing(data2019_split)

data2019_train <- data2019_train %>% mutate(
  social_support = `Social support`,
  freed_choices = `Freedom to make life choices`,
  generosity = `Generosity`,
  perc_corrup = `Perceptions of corruption`,
  health_life = `Healthy life expectancy`
)

data2019_test <- data2019_test %>% mutate(
  social_support = `Social support`,
  freed_choices = `Freedom to make life choices`,
  generosity = `Generosity`,
  perc_corrup = `Perceptions of corruption`,
  health_life = `Healthy life expectancy`
)

```

LINEAR REGRESSIN MODEL 1: FREEDOM TO MAKE LIFE CHOICES

```{r}

mod1 <- lm(`Happiness score` ~ `Freedom to make life choices`,
                        data = data2019_train)
summary(mod1)

ggplot(mod1, aes(x = `Freedom to make life choices`, y = `Happiness score`)) +
  geom_point() + geom_smooth(method = "lm") +
  labs(title = "Freedom to Make Life Choices on Happiness Score (2019)")

```


LOOKING AT VARIABLE IMPORTANCE

```{r}

rf_fit3 <- randomForest(
  `Happiness score` ~ social_support + freed_choices + generosity + perc_corrup +
    health_life, data = data2019_train)
print(rf_fit3)

importance(rf_fit3)

```

```{r}

varImpPlot(rf_fit3)

```

LINEAR REGRESSION MODEL 2: SOCIAL SUPPORT

```{r}

mod2 <- lm(`Happiness score` ~ `Social support`,
           data = data2019_train)
summary(mod2)

ggplot(mod2, aes(x = `Social support`, y = `Happiness score`)) +
  geom_point() + geom_smooth(method = "lm")+
  labs(title = "Social on Happiness Score (2019)")

```



# PREDICTED TRUE PLOTS: FREEDOM TO MAKE LIFE CHOICES AND SOCIAL SUPPORT

```{r}

mod3 <- lm(`Happiness score` ~ freed_choices + social_support,
           data = data2019_train)
summary(mod3)

```

```{r}

preds_train <- predict(mod3, newdata = data2019_train)
preds_test <- predict(mod3, newdata = data2019_test)

results_train <- 
  tibble(
    `preds` = preds_train,
    `true` = paste(data2019_train$freed_choices, data2019_train$social_support),
    `type` = "train"
  )

results_test <- 
  tibble(
    `preds` = preds_test,
    `true` = paste(data2019_test$freed_choices, data2019_test$social_support),
    `type` = "test"
  )

results_df <- 
  bind_rows(results_train, results_test)

ggplot(results_df, aes(x = true, y = preds)) +
  geom_point(aes(color = type)) + 
  geom_abline(color = "red") +
  facet_wrap(~ type) +
  ylim(0,10) +
  theme_minimal(base_size = 16) + 
  theme(legend.position="bottom") +
  labs(title = "Predicted Trues: Freedom & Social Support (2019)")

```

CREATING RIDGE REGRESSION & LOOKING AT COEFFICIENTS

```{r}

ridge_mod1 <- cv.glmnet(`Happiness score` ~ social_support + freed_choices,
                       data = data2019_train,
                       alpha = 0)
plot(ridge_mod1)

```

```{r}

coef(ridge_mod1, s = ridge_mod$lambda.min)

```



## DATA 2020

THIS CODE DEMONSTRATES DATA WITH RECORDED HAPPINESS SCORES AS IS

```{r}

data_happy7 <- data2020 %>% filter(`Happiness score` >= 7.0) %>% select(Country | `Happiness score`)
print(data_happy7)

```


```{r}

ggplot(data_happy7, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Countries w/Happiness Score >= 7 (2020)")

```


```{r}

data_happy6 <- data2020 %>% filter(`Happiness score` >= 6.0, `Happiness score` < 7.0) %>% 
  select(Country | `Happiness score`)
print(data_happy6)

```


```{r}

ggplot(data_happy6, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 6 & < 7 (2020)")

```


```{r}

data_happy5 <- data2020 %>% filter(`Happiness score` >= 5.0, `Happiness score` < 6.0) %>% 
  select(Country | `Happiness score`)
print(data_happy5)

```


```{r}

ggplot(data_happy5, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 5 & < 6 (2020)")

```


```{r}

data_happy4 <- data2020 %>% filter(`Happiness score` >= 4.0, `Happiness score` < 5.0) %>% 
  select(Country | `Happiness score`)
print(data_happy4)

```


```{r}

ggplot(data_happy4, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 4 & < 5 (2020)")

```


```{r}

data_happy3 <- data2020 %>% filter(`Happiness score` >= 3.0, `Happiness score` < 4.0) %>% 
  select(Country | `Happiness score`)
print(data_happy3)

```


```{r}

ggplot(data_happy3, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 3 & < 4 (2020)")

```


```{r}

data_happy2 <- data2020 %>% filter(`Happiness score` >= 2.0, `Happiness score` < 3.0) %>% 
  select(Country | `Happiness score`)
print(data_happy2)

```


```{r}

ggplot(data_happy2, aes(x = Country, y = `Happiness score`, fill = `Happiness score`)) +
  geom_bar(stat = "identity", position = "dodge") + ylim(0,8) +
  labs(title = "Countries' Happiness Score >= 2 & < 3 (2020)")

```


DESCRIPTIVE PLOT AS SCATTER PLOTS

```{r}

data_list <- list(`Scores 7+` = data_happy7, `7 > Scores > 6` = data_happy6, `6 > Scores > 5` = data_happy5,
                  `4 > Scores > 5` = data_happy4, `3 > Scores > 4` = data_happy3, `2 > Scores > 3` = data_happy2)
combined_data <- bind_rows(data_list, .id = "dataset")
ggplot(combined_data, aes(x = Country, y = `Happiness score`)) +
  geom_point() +
  labs(title = "Happiness scores 2020", x = "Country", y = "Happiness Score (out of 10)") +
  facet_wrap(~ dataset)

```

SPLITTING THE DATA INTO TRAIN & TEST SETS

```{r}

data2020_split <- initial_split(data2020, prop = 0.75)
data2020_train <- training(data2020_split)
data2020_test <- testing(data2020_split)

data2020_train <- data2020_train %>% mutate(
  social_support = `Social support`,
  freed_choices = `Freedom to make life choices`,
  generosity = `Generosity`,
  perc_corrup = `Perceptions of corruption`,
  health_life = `Healthy life expectancy`
)

data2020_test <- data2020_test %>% mutate(
  social_support = `Social support`,
  freed_choices = `Freedom to make life choices`,
  generosity = `Generosity`,
  perc_corrup = `Perceptions of corruption`,
  health_life = `Healthy life expectancy`
)

```


LINEAR REGRESSION MODEL 4: FREEDOM TO MAKE LIFE CHOICES

```{r}

mod4 <- lm(`Happiness score` ~ freed_choices,
           data = data2020_train)
summary(mod4)

```


```{r}

ggplot(mod4, aes(x = freed_choices, y = `Happiness score`)) +
  geom_point() + geom_smooth(method = "lm") +
  labs(title = "Freedom to Make Life Choices on Happiness Score (2020)")

```


LOOKING AT VARIABLE IMPORTANCE

```{r}

rf_fit2 <- randomForest(
  `Happiness score` ~ social_support + freed_choices + generosity + perc_corrup +
    health_life, data = data2020_train)

importance(rf_fit2)

```

```{r}

varImpPlot(rf_fit2)

```


LINEAR REGRESSION MODEL 5: SOCIAL SUPPORT

```{r}

mod5 <- lm(`Happiness score` ~ `Social support`,
           data = data2020_train)
summary(mod5)

```


```{r}

ggplot(mod5, aes(x = `Social support`, y = `Happiness score`)) +
  geom_point() + geom_smooth(method = "lm") +
  labs(title = "Social Support on Happiness Score (2020)")


```


PREDICTED TRUE PLOTS

```{r}

preds_train <- predict(mod, newdata = data2020_train)
preds_test <- predict(mod, newdata = data2020_test)

results_train <- 
  tibble(
    `preds` = preds_train,
    `true` = paste(data2020_train$freed_choices, data2020_train$social_support),
    `type` = "train"
  )

results_test <- 
  tibble(
    `preds` = preds_test,
    `true` = paste(data2020_test$freed_choices, data2020_test$social_support),
    `type` = "test"
  )

results_df <- 
  bind_rows(results_train, results_test)

ggplot(results_df, aes(x = true, y = preds)) +
  geom_point(aes(color = type)) + 
  geom_abline(color = "red") +
  facet_wrap(~ type) +
  theme_minimal(base_size = 16) + 
  theme(legend.position="bottom") +
  labs(title = "Predicted Trues: Freedom & Social Support (2020)")

```


RIDGE REGRESSION

```{r}

ridge_mod2 <- cv.glmnet(`Happiness score` ~ social_support + freed_choices,
                        data = data2020_train,
                        alpha = 0)
plot(ridge_mod2)

```

```{r}

coef(ridge_mod2, s = ridge_mod$lambda.min)

```

